\setlength{\parindent}{15pt}
\setlength{\parskip}{0em}

\chapter*{Acknowledgements}
\vspace{-15mm}

Everything started while I was doing my internship for my masters thesis in Paris at Aldebaran Robotics and a colleague sent to me the link of opening PhD positions at the CVC. I initially doubted whether to apply since I always considered myself to have a more engineer background and at that time was not really aware about the consequences of starting a PhD and how would affect to my career in the future.

I am an outgoing person which like challenges. This is the reason why I decided to send an email to just ask information about the different open positions and arrange a call meeting with Dani. That was my first contact with Dani, which later has become my PhD supervisor during this last 5 years. I would like to thank him for giving me this opportunity and the support for all the collaborations and internships these years.

My stay at the CVC was very productive and gave me the opportunity to meet amazing people such as Felipe, to whom I want to thank for the enormous number of hours of discussion about redundant questions for geometry. The two Paus, Pep, Albert, Xavi and Arnau for the good times spent in our cluster office in the basement; Lorena, Iris, Xesco, Marc, Felipe, Bojana and the rest of the CARLA team for the nice breakfasts and barbeque; Alexandra, Txell and the administration people to help and listen me every time I wanted to organise hackathons or promote any crazy open source initiative.

Probably I am missing a long list of people that I met during these years since the number of collaborations and contacts I did has been remarkable. However, I would specially like to thank Vassilis and Krystian to give me the opportunity to start collaborating with them and guide me during my initial stage of the PhD. The rest of the Imperial College people, Axel, Adri, Joan, Marina, Irene, Pablo, Yurun for such good times in London and China.

My tight collaborations with the open source community brought me also the opportunity to meet brilliant people like Gary and Ethan; Vincent, Vadim, Stefano, Reza, Mona, Wes, Prassanna from OpenCV and Arraiy. During my stay in the US, Camilo, Karsten, Ali, Matilda and Anguelos for such good times and never ending discussions about open source and computer science. I would also specially thank to Francesc, Luis, Antonio, Albert, Jordi and the rest of people from Kognia Sports and IRI for helping me during my last stage of the PhD.

Finally, to Gore and Daga for the initial support to this long trip and bring inspiration and meaning to my work. A strong mention to my family for trying to understand my going and coming back trips, and not to forgive the ones that could not make it to see the final result of this thesis. 

\setlength{\parskip}{1em}
